---
title: "Juan Carlos Abiotic Data for Prather et al 2019"
author: "Matthew Chmielewski"
date: "May 8, 2019"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

Load data and libraries
```{r}
hoboallyears<-read.csv("JC_Hobo_Master.csv")
library(lubridate)
library(ggplot2)
library(tidyverse)
library(plyr)
library(ggridges)
```

Check for NA values, create time and date variables
```{r}

head(hoboallyears)
length(hoboallyears$Date.Time)
#drop na values
hoboallyearsa<-na.omit(hoboallyears)
length(hoboallyears$Date.Time)
#looks like there were no NA values
#create wanted time variables via lubridate
hoboallyearsa$day<-day(mdy_hm(hoboallyearsa$Date.Time))
hoboallyearsa$month<-month(mdy_hm(hoboallyearsa$Date.Time), label = TRUE)
hoboallyearsa$year<-year(mdy_hm(hoboallyearsa$Date.Time))
hoboallyearsa$date<-as.character(paste(hoboallyearsa$month, hoboallyearsa$day, sep = ""))
hoboallyearsa$time<-paste(c(hour(mdy_hm(hoboallyearsa$Date.Time))), minute(mdy_hm(hoboallyearsa$Date.Time)), sep=":")
hoboallyearsa$julian<-yday(date(mdy_hm(hoboallyearsa$Date.Time)))
hoboallyearsa$monthyear<-as.factor(paste(hoboallyearsa$month, hoboallyearsa$year, sep = "."))
```

The 2013 data came in two formats that are in a different (dmy) order
We can split data into already usable data as well as the two sections that need to be fixed, process the 2013 data via lubridate, then add it back
to the rest of the dataset
```{r}
american2013<-subset(hoboallyearsa, date == "NANA")
chilean2013<-subset(hoboallyearsa, year == "2013")
restofdata<-subset(hoboallyearsa, year != "2013" & date != "NANA")
length(american2013[,1])+length(chilean2013[,1])+length(restofdata[,1])
head(american2013)

#convert chilean (dmy) dates
chilean2013$day<-day(dmy_hm(chilean2013$Date.Time))
chilean2013$month<-month(dmy_hm(chilean2013$Date.Time), label = TRUE)
chilean2013$year<-year(dmy_hm(chilean2013$Date.Time))
chilean2013$date<-as.character(paste(chilean2013$month, chilean2013$day, sep = ""))
chilean2013$time<-paste(c(hour(dmy_hm(chilean2013$Date.Time))), minute(dmy_hm(chilean2013$Date.Time)), sep=":")
chilean2013$julian<-yday(date(dmy_hm(chilean2013$Date.Time)))
chilean2013$monthyear<-as.factor(paste(chilean2013$month, chilean2013$year, sep = "."))

#should be able to process via lubridate now
american2013$day<-day(mdy_hms(american2013$Date.Time))
american2013$month<-month(mdy_hms(american2013$Date.Time), label = TRUE)
american2013$year<-year(mdy_hms(american2013$Date.Time))
american2013$date<-as.character(paste(american2013$month, american2013$day, sep = ""))
american2013$time<-paste(c(hour(mdy_hms(american2013$Date.Time))), minute(mdy_hms(american2013$Date.Time)), sep=":")
american2013$julian<-yday(date(mdy_hms(american2013$Date.Time)))
american2013$monthyear<-as.factor(paste(american2013$month, american2013$year, sep = "."))


#now need to stitch together "american2013", "chilean2013", and "restofdata"

hobosfullset<-full_join(restofdata, chilean2013)
hobosfullset<-full_join(hobosfullset, american2013)
length(hobosfullset)
```

In order to properly process the dataset we have to convert some columns to factors and properly temporally order monthyear
```{r}
levels(hobosfullset$monthyear)

#yearmonth is not a factor, need to coerce it into factor format
hobosfullset$monthyear<-as.factor(hobosfullset$monthyear)
levels(hobosfullset$monthyear)

#now monthyear is a factor but levels are not in temporal order, need to put them in order

hobosfullset$monthyear<-factor(hobosfullset$monthyear, levels = c("Jan.2013", "Feb.2013", "Mar.2013", "Apr.2013", "May.2013", "Jun.2013", "Jul.2013", "Jan.2014", "Feb.2014", "Mar.2014", "Apr.2014", "May.2014", "Feb.2015", "Mar.2015", 
                                                                   "Apr.2015", "May.2015", "Jun.2015", "Jul.2015", "Aug.2015", "Sep.2015", "Oct.2015", "Nov.2015", "Dec.2015", "Jan.2016", "Feb.2016")) 

levels(hobosfullset$monthyear)

#hobosfullset$monthyear, gsub(":[[:digit:]]+", "", hobosfullset$time)

#now in order!
```

Now we can summarise data properly
```{r}
#summarise by month
sumjchobos<- ddply(hobosfullset, c("monthyear","Treatment"), summarise,
                         NTemp  = length(Temp),
                         meanTemp = mean(Temp),
                         sdTemp   = sd(Temp),
                         seTemp = sdTemp / sqrt(NTemp),
                         min = min(Temp),
                         max = max(Temp),
                        
                        NRH  = length(RH),
                        meanRH = mean(RH),
                        sdRH   = sd(RH),
                        seRH = sdRH / sqrt(NRH),
                        min = min(RH),
                        max = max(RH)
                        
                       
)
sumjchobos$CITemp<-sumjchobos$seTemp*1.96
sumjchobos$CIRH<-sumjchobos$seRH*1.96

head(sumjchobos)

write.csv(sumjchobos, "sumjchobos.csv")
```

Now we can plot abiotic data by the timeframe we are interested in
```{r}
#plotting humidity
hoboshumidity<-ggplot(data = hobosfullset, aes(monthyear, RH, fill = Treatment))+
  geom_boxplot() + 
  ggtitle("Humidity")+
theme(#axis.title.x = element_blank(),
  #axis.title.y = element_blank(),
  axis.text.x  = element_text(angle = 90, vjust=1, size=12),
  panel.grid.minor.x=element_blank(),
  panel.grid.major.x=element_blank(),
  panel.grid.minor.y=element_blank(),
  panel.grid.major.y=element_blank())
  
hoboshumidity  

hobostemp<-ggplot(data = hobosfullset, aes(monthyear, Temp, fill = Treatment))+
  geom_boxplot()+  
  ggtitle("Temp")+
  theme(#axis.title.x = element_blank(),
    #axis.title.y = element_blank(),
    axis.text.x  = element_text(angle = 90, vjust=1, size=12),
    panel.grid.minor.x=element_blank(),
    panel.grid.major.x=element_blank(),
    panel.grid.minor.y=element_blank(),
    panel.grid.major.y=element_blank())

hobostemp 

pdf(file="hoboshumidity.pdf", width = 12, height = 11, family = "Helvetica")
hoboshumidity
dev.off()

pdf(file="hobostemp.pdf",width=12, height=11, family = "Helvetica")
hobostemp
dev.off()

```

Ridgeplots of temp and temperature differential
```{r}

hobostempridges<-ggplot(data = subset(hobosfullset, year != "2013"), aes(x=Temp, y=monthyear, fill=Treatment))+
  geom_density_ridges(scale=4)

hobostempridges

pdf(file="hobostempridges.pdf", width =12, height=11, family = "Helvetica")
hobostempridges
dev.off()

#what about a ridgeplot of difference values (OTC-CON across time, positive values indicate heating, negative indicate cooler than control)
#minutes and seconds might be off, but should be able to do this by hour of the day 

differentialdata<-hobosfullset
differentialdata$hour<-gsub(":[[:digit:]]+", "", differentialdata$time)
differentialdata$hour

differentialdata<-unite(differentialdata, datehour, monthyear, hour, remove=FALSE)
head(differentialdata)

differentialdata1<-subset(differentialdata, year != "2013")




sumdifferential<- ddply(differentialdata1, c("datehour","Treatment"), summarise,
                   NTemp  = length(Temp),
                   meanTemp = mean(Temp),
                   sdTemp   = sd(Temp),
                   seTemp = sdTemp / sqrt(NTemp),
                   min = min(Temp),
                   max = max(Temp),
                   
                   NRH  = length(RH),
                   meanRH = mean(RH),
                   sdRH   = sd(RH),
                   seRH = sdRH / sqrt(NRH),
                   min = min(RH),
                   max = max(RH)
                   
                   
)

head(sumdifferential)

diffhours<-unique(sumdifferential$datehour)
diffmonths<-gsub("_[[:digit:]]+", "", diffhours)
diffmonths

cont<-grep("CON", sumdifferential$Treatment)
otc<-grep("OTC", sumdifferential$Treatment)

difftemps<-sumdifferential$meanTemp[otc]-sumdifferential$meanTemp[cont]
differentials<-data.frame(diffhours, difftemps, diffmonths)

differentials$diffmonths<-factor(differentials$diffmonths, levels = c("Jan.2014", "Feb.2014", "Mar.2014", "Apr.2014", "May.2014", "Feb.2015", "Mar.2015", 
                                                                  "Apr.2015", "May.2015", "Jun.2015", "Jul.2015", "Aug.2015", "Sep.2015", "Oct.2015", "Nov.2015", "Dec.2015", "Jan.2016", "Feb.2016")) 



hobosdiffridges<-ggplot(data = differentials, aes(x=difftemps, y=diffmonths))+
  geom_density_ridges(scale=4)+
  geom_vline(xintercept = 0, size= 1, linetype= 2)+
  theme_bw()+
  theme(panel.grid.minor.y=element_blank(),
           panel.grid.major.y=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.x=element_blank())+
  xlab("Temperature Differential")+
  ylab("")
  

hobosdiffridges

pdf(file="hobosdiffridges.pdf", width =12, height=11, family = "Helvetica")
hobosdiffridges
dev.off()

png(file="hobosdiffridgespnd.png", width=1150, height=1150, res=150)
hobosdiffridges
dev.off()
```
